An Analysis of Wind Energy Potential in the North Sea

### 2024-01-29 paper/paper.tex
fix typos

%For assessing the viability of wind turbine (parks), one might also be interested in the weekly or daily values  and, even more importantly, in a forecast. Hence, for interpolating the daily expected power density, we employ a Gaussian Process Regression model to perform a time-series regression. In addition to interpolating, we also use the same model for extrapolation and thereby making predictions about the future daily expected power density.
The first step for conducting a GP regression is to decide on a kernel $k$ for the prior
\eqref{eq:gp_prior}. By analysing the plotted raw monthly expected power density, we identify a
yearly periodic trend and a slight global downward trend. The former can be captured by the
Exponential-Sine-Squared kernel (ES) \cite{MacKay1998IntroductionTG}. Supported by the empirical
data, we choose a periodicity of $365 \, \text{days}$ and an amplitude of $350 \,
\frac{\textrm{W}}{\text{m}^2}$. For the length scale, we settle for a rather small value ($< 2 \, $
days) in order to capture local patterns in the periodicity more accurately. For allowing the
amplitude to slowly decay over long periods of time, we multiply the ES kernel by a very smooth
(length scale $> 50 \, \text{years}$) Squared-Exponential (SE) kernel. For reproducing the global
downwards trend observed in section \ref{sec:trend}, we also choose a very smooth
Squared-Exponential kernel with an output scale of $100 \, \frac{\textrm{W}}{\text{m}^2}$. For
accounting for the monthly midterm trend, we add a Rational-Quadratic kernel
\cite{rasmussen-williams-gp} with a scale of $10 \, \frac{\textrm{W}}{\text{m}^2}$ and a length
scale of one month. Finally, the additive kernel, consisting of a White-Noise (WN) and a Matern
kernel (MA) \cite{abramowitz1968handbook}, is responsible for explaining measurement noise in the
data and thus also prevents overfitting. Thereby, the kernel has the form 
$$k = \underbrace{\text{ES} \cdot \text{SE}}_{\text{yearly trend}} 
+ \underbrace{\text{RQ}}_\text{monthly trend} 
+ \underbrace{\text{ES}}_\text{longterm trend} 
+ \underbrace{\text{WN} 
+ \text{MA}}_{\text{noise}}.$$
By drawing samples\footnote{with the mean of the data added.} $f_i \sim
\mathcal{GP}(f; 0, k)$ from the prior distribution \eqref{eq:gp_prior}, we gauge how accurately our
kernel choice fits to the given data. The results are shown in Figure \ref{fig:gp-samples}. We can
observe that the amplitude and periodicity of some of the samples mirrors the data's in an accurate
manner, supporting our choice of hyperparameters. After conditioning the Gaussian process prior on
the normalised data, we obtain the posterior distribution $\mathcal{GP}(f; m', k')$
\eqref{eq:gp_posterior}. The evaluated posterior mean and standard deviation can be found in Figure
\ref{fig:gp-pred}. We can observe that the mean function is following the periodicity of the ground
truth without overfitting extreme power densities. Furthermore, the uncertainty of the model
(standard deviation) increases gradually the further the prediction is from the data. Notably, the
model also captures the global downward trend. When comparing the prediction to the test data in the
years 2017 and 2018, we can observe that the prediction fits the ground truth accurately and all
data points are well within one standard deviation of the mean.

@@

However, there are some shortcomings. One of the inherent limitations of the GPR is that it yields
functions mapping onto $\R$, while the power density is only limited to $\R_+$.
%Further, we conducted an interpolation of daily values based on monthly data, which does not take the inter-day variability into account. For the purpose of extrapolating daily power densities, 
%it might be advisable in further work to rely on daily data to begin with, complicating the choice of the kernel.  
Another important limitation is that our data was recorded at a height of $4.38 \mathrm{m}$, whereas
the real height of wind turbines can exceed $100 \mathrm{m}$. One commonly applied rule of thumb to
mitigate this issue is multiplying the wind speeds by the factor $\left(\frac{h}{4.38 \mathrm{m}}
\right)^{0.14}$ or the power density by its cube, where $h$ is the  proposed hub height of a turbine
\citep{statanalysis}. However, this is far from exact, as the winds at different heights could be
subject to different circulations.  Furthermore, we did not take the direction of the wind into
account and made the simplification that wind turbines are always perfectly aligned.  Also, for
equation \eqref{eq:power}, we assumed a constant air pressure rather than considering the monthly
changing pressure.  Lastly, with the use of the maximum-likelihood method, we assumed the
independence of measurements. While this is generally not given, we were still able to produce
adequate estimations.

>>>>

%For assessing the viability of wind turbine (parks), one might also be interested in the weekly or daily values  and, even more importantly, in a forecast. Hence, for interpolating the daily expected power density, we employ a Gaussian Process Regression model to perform a time-series regression. In addition to interpolating, we also use the same model for extrapolation and thereby making predictions about the future daily expected power density.
The first step for conducting a GP regression is to decide on a kernel $k$ for the prior
\eqref{eq:gp_prior}. By analysing the plotted raw monthly expected power density, we identify a
yearly periodic trend and a slight global downward trend. The former can be captured by the
Exponential-Sine-Squared kernel (ES) \cite{MacKay1998IntroductionTG}. Supported by the empirical
data, we choose a periodicity of $365 \, \text{days}$ and an amplitude of $350 \,
\frac{\textrm{W}}{\text{m}^2}$. For the length scale, we settle for a rather small value ($< 2 \, $
days) in order to capture local patterns in the periodicity more accurately. For allowing the
amplitude to slowly decay over long periods of time, we multiply the ES kernel by a very smooth
(length scale $> 50 \, \text{years}$) Squared-Exponential (SE) kernel. For reproducing the global
downwards trend observed in section \ref{sec:trend}, we also choose a very smooth
Squared-Exponential kernel with an output scale of $100 \, \frac{\textrm{W}}{\text{m}^2}$. For
accounting for the monthly midterm trend, we add a Rational-Quadratic kernel
\cite{rasmussen-williams-gp} with a scale of $10 \, \frac{\textrm{W}}{\text{m}^2}$ and a length
scale of one month. Finally, the additive kernel, consisting of a White-Noise (WN) and a Matern
kernel (MA) \cite{abramowitz1968handbook}, is responsible for explaining measurement noise in the
data and thus also prevents overfitting. Thereby, the kernel has the form 
$$k = \underbrace{\text{ES} \cdot \text{SE}}_{\text{yearly trend}} 
+ \underbrace{\text{RQ}}_\text{monthly trend} 
+ \underbrace{\text{ES}}_\text{longterm trend} 
+ \underbrace{\text{WN} 
+ \text{MA}}_{\text{noise}}.$$
By drawing samples\footnote{with the mean of the data added.} $f_i \sim
\mathcal{GP}(f; 0, k)$ from the prior distribution \eqref{eq:gp_prior}, we gauge how accurately our
kernel choice fits to the given data. The results are shown in Figure \ref{fig:gp-samples}. We can
observe that the amplitude and periodicity of some of the samples mirrors the data's in an accurate
manner, supporting our choice of hyperparameters. After conditioning the Gaussian process prior on
the normalised data, we obtain the posterior distribution $\mathcal{GP}(f; m', k')$
\eqref{eq:gp_posterior}. The evaluated posterior mean and standard deviation can be found in Figure
\ref{fig:gp-pred}. We can observe that the mean function is following the periodicity of the ground
truth without overfitting extreme power densities. Furthermore, the uncertainty of the model
(standard deviation) increases gradually the further the prediction is from the data. Notably, the
model also captures the global downward trend. When comparing the prediction to the test data in the
years 2017 and 2018, we can observe that the prediction fits the ground truth accurately and almost all
data points are well within one standard deviation of the mean.

@@

However, there are some shortcomings. One of the inherent limitations of the GPR is that it yields
functions mapping onto $\R$, while the power density is only limited to $\R_+$.
%Further, we conducted an interpolation of daily values based on monthly data, which does not take the inter-day variability into account. For the purpose of extrapolating daily power densities, 
%it might be advisable in further work to rely on daily data to begin with, complicating the choice of the kernel.  
Another important limitation is that our data was recorded at a height of $4.38 \mathrm{m}$, whereas
the real height of wind turbines can exceed $100 \mathrm{m}$. One commonly applied rule of thumb to
mitigate this issue is multiplying the wind speeds by the factor $\left(\frac{h}{4.38 \mathrm{m}}
\right)^{0.14}$ or the power density by its cube, where $h$ is the  proposed hub height of a turbine
\citep{statanalysis}. However, this is far from exact, as the winds at different heights could be
subject to different circulations.  Furthermore, we did not take the direction of the wind into
account and made the simplification that wind turbines are always perfectly aligned.  Also, for
equation \eqref{eq:power}, we assumed a constant air density $\rho$ rather than considering the monthly
changing densities.  Lastly, with the use of the maximum-likelihood method, we assumed the
independence of measurements. While this is generally not given, we were still able to produce
adequate estimations.
